### 2021-08-21 tSNE Introduction Document.Rmd

---
title: "TSNE Introduction"
author: "Paul Harmon"
date: "6/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rtsne)
```

# Introduction

This chapter should act as a pseudo-paper that breaks out the derviation and mathematical details of t-SNE in detail. Additionally, should overview the pros and cons of the method and discuss how it should be used.

This will act as an addendum to my introduction section but should also be submittable as a stand-alone paper.

## How is t-SNE used

## Comparison methods

### SNE
### MDS/Sammon Mapping
### Umap

# t-SNE Algorithm and Math

## Derivation

## Mathematical Details

## PCA as Initial Step

## Considerations

# Perplexity

## Cost Decreases as Function of Perplexity

## Perplexity