### 2021-09-21 Simulation Study Document V4.Rmd

---
title: "Simulation Study Document V4"
author: "Paul Harmon"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

source("OutlierCompNew.R")
library(ggplot2)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(janitor)
library(magrittr)
library(knitr)
library(Rtsne)
library(robustHD)
library(GGally)
library(stringr)
library(plotly)
library(R.utils)
library(vegan)
library(mnormt)

```

# Introduction

This doc contains simulation studies around the efficacy of a Permanova-based method for identifying influential points in a t-SNE or MDS map.  The document steps through a process of simulating data from 3 different groups (currently, these groups are clearly separated) with *3* influential points simulated on a standardized scale 5-6 standard deviations above the observed data.

The code below outlines the number of simulations run - currently this number is small for computational ease but will be increased (with the use of adaptive Permanova methods).

```{r, echo = FALSE, eval = TRUE}

n_sims = 100

#create a bunch of lists to store stuff later
plot_data_list <- list()
plot_tsne_results_list <- list()
plot_mds_results_list <- list()
plot_tsne_fstat_results_list <- list()
plot_mds_fstat_results_list <- list()
results_list <- list()
results_list_cmd <- list()
df3list = df2list = list()

```

The below chunk runs the simulations. Each iteration simulates a new (contaminated) dataset, runs the t-SNE and MDS mapping procedure, and tests the consistency of the configurations for each of the *n* holdout points. **This code is not run in this version.**

Several versions of the simulations will be run:

+ 1 group with 3 "influental" points
+ 3 groups with 3 "influential" points with different variances
+ 3 groups with 3 influential points with same variances

+ Sample size increase/decrease (number of influential points relative to sample size too)

## Simulation Chunks

This chunk runs the simulation with slighlty increased variability in the three groups, and a restructuring of the influential points. Rather than simulating as outliers all on the positive side, I simulate one that has a large positive set of outlying points, one with a large negative set of outlying points, and one that simply 'bounces around' the multivariate space and breaks the grouping structure.

```{r sims_v2, eval = TRUE, cache = TRUE, warning = FALSE, message = FALSE}
perplexity = 10

for(j in 1:n_sims){

  #### Simulate the Data with outliers and a single group (simulates all from one group with mean 40)

  g1 <- rmnorm(n = 10, mean = rep(40, 10), diag(45, 10)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(40, 20), diag(45, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(40, 30), diag(45, 30)) %>%
    t() %>%
    data.frame()

  yes_structure <- rbind(g1, g2, g3) %>%
    data.frame() %>%
    mutate(Group = c(rep("1", 10), rep("2", 20), rep("3", 30)))

  yes_structure_scale <- lapply(yes_structure[,1:10], scale) %>% as_tibble()

  ### adds outliers - but not as a single group. Here we have an outlier that includes a really big one, a really small observation, and a very abnormal one
  add_outliers <- rbind(rnorm(10, 3, .5), rnorm(10, -3, .5), rnorm(10, c(rep(-2,3),rep(0,3),rep(2,3),0), .5)) %>% as_tibble()
  names(add_outliers) <- names(yes_structure_scale)

  yes_structure_final <- bind_rows(yes_structure_scale, add_outliers) 
  yes_structure_final$Group <- c(yes_structure$Group, rep("Outlier",3))

  ##### Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:10) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
  plot_data_list[[j]] <- datplot

  #### Run the Permanova mapping (for t-SNE)

  xTSNE <- MultiPermanova(yes_structure_final[,1:10], maptype = "tSNE")
  results_list[[j]] <- xTSNE

  df2 = tibble(Holdout = 1:length(xTSNE$modellist), PValues = sapply(xTSNE$modellist, pullPval), Fstat = sapply(xTSNE$modellist, pullFstat))
  df2$Colors <- ifelse(df2$PValues < 0.05, TRUE, FALSE)
  df2list[[j]] = df2

  #### TSNE Plots########################################################
  #Visualize the results and store in a list
  sigplot <- ggplot(df2, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: P-Values From Adonis")
  plot_tsne_results_list[[j]] <- sigplot

  #Visualize the f stats and store in a list
  sigplotf <- ggplot(df2, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: F Stats From Adonis")
  plot_tsne_fstat_results_list[[j]] <- sigplotf

   #### Run the Permanova mapping (for MDS)

   xCMD <- MultiPermanova(yes_structure_final[,1:10], maptype = "MDS")
  results_list_cmd[[j]] <- xCMD

  ### CMD Plots #########################################################

  df3 = tibble(Holdout = 1:length(xCMD$modellist), PValues = sapply(xCMD$modellist, pullPval), Fstat = sapply(xCMD$modellist, pullFstat))
    df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)
  df3list[[j]] = df3

  #Visualize the results and store in a list
  sigplot2 <- ggplot(df3, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: P-Values From Adonis")
  plot_mds_results_list[[j]] <- sigplot2

  #Visualize the f stats and store in a list
  sigplot2f <- ggplot(df3, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: F Stats From Adonis")
  plot_mds_fstat_results_list[[j]] <- sigplot2f

}
```

# Results {.tabset}

The simulated datasets can be seen in the tabset below.

## Plot List
```{r}
library(ggpubr)
#ggarrange(plotlist = plot_data_list)
print(plot_data_list)
```

## Summary Metrics

We can start to look at some results below. The code generates the rate of correctly classified influential points for both the MDS (shown first) and t-SNE maps. The MDS correctly classifies the influential point *some* of the time, but the t-SNE method very rarely classifies the influential points as having an impact on the configuration.

```{r}

#print the head of the df lists
df3list %>% head(3)
### Identify the number of CORRECTLY Classified influential points
#extract the p-values from the last 3 from MDS
influential_correct <- function(temp){length(which(tail(temp, 3)$PValues < 0.05))/3}
mds_correct <- lapply(df3list, influential_correct)

# extract the pvalues from the last 3 of the t-SNE
tsne_correct <- lapply(df2list, influential_correct)

dat = tibble(MDS = unlist(mds_correct), TSNE = unlist(tsne_correct))
#plot(dat$MDS, dat$TSNE)

mds_correct %>% unlist() %>% mean() %>% kable(col.names = "MDS Percent Correctly Classified")
tsne_correct %>% unlist() %>% mean() %>% kable(col.names = "t-SNE Percent Correctly Classified")

```

Now we can look at the rate of misclassified non-influential points. These are the 60 'regular' points that are spuriously detected as outliers. Each percentage is calculated as a percentage of the 60 that are NOT small p-values (currenlty < 0.05 without a correction). We want a high percentage of correctly-identified non-influential points (although given the machinery of the hypothesis tests used, we would expect at least some spurious detections).

Both are relatively high percentages (as we would expect) - but the t-SNE hovers right around the expected 0.95 and MDS is higher.

```{r}

### Identify rate of CORRECTLY Classified non-influential points
#extract the p-values from the last 3 from MDS
ni_correct <- function(temp){length(which(head(temp, 60)$PValues >= 0.05))/60}

correct_mds <- lapply(df3list, ni_correct)
#kable(correct_mds, col.names = "Correctly Identified Non-Influential: MDS")

# extract the pvalues from the last 3 of the t-SNE
correct_tsne <- lapply(df2list, ni_correct)
#kable(correct_tsne, col.names = "Correctly Identified Non-Influential: tSNE")

correct_mds %>% unlist() %>% mean() %>% kable(col.names = "MDS Percent Correctly Classified Non-Influential")
correct_tsne %>% unlist() %>% mean() %>% kable(col.names = "t-SNE Percent Correctly Classified Non-Influential")

```

# Plots {.tabset}

## MDS P Values

Plots are shown below for the MDS .

```{r}
plot_mds_results_list
```

## MDS F Stats

```{r}
plot_mds_fstat_results_list
```

## tSNE

```{r}
plot_tsne_results_list
```

## t-SNE F Stats

```{r}
plot_tsne_fstat_results_list
```