### 2021-09-21 Test.Rmd

---
title: "Python Methods for Skillz App"
author: "Paul Harmon"
date: "8/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reads in the python files that we'll use.

```{python}
import numpy as np
import pandas as pd

peopleSkillz = pd.read_csv("https://raw.githubusercontent.com/AtriumAI/InternalSkillzApp/Keely26/issue138/temp_data%20/peopleSkillz.csv?token=AFIT5TOPNWUQ7MJN2ZYIGO3BGKWAC")
roleSkillz = pd.read_csv("https://raw.githubusercontent.com/AtriumAI/InternalSkillzApp/Keely26/issue138/temp_data%20/roleSkillz.csv?token=AFIT5TJ4TKCRILP5FNP7LX3BGKWEC")

peopleSkillz.head()
roleSkillz.head()
```

```{python}
prankings = pd.read_csv("C:/Users/paulh/Downloads/people_rankings.csv")
roles_temp = pd.read_csv("C:/Users/paulh/Downloads/people_rankings.csv")

roles_temp.head()
```

```{python}
#skill: array of skills for a given person
#role: array of roles for a given person
def match_skillz(persondf, roledf): 
  temp = persondf >= role 
  pct = sum(temp)/temp.size
  return pct
```

```{python}
#Method to generate a comparison between two vectors

#temp = np.array([3,3,2])  np.array([3,4,3])
#sum(temp)/temp.size

#peopleSkillz.columns.values.tolist()[1:]
#roleSkillz.columns.values.tolist()[3:]

#peopleSkillz[peopleSkillz.columns.values.tolist()[:]]
peopleSkillz[peopleSkillz['full_name'] == "Weisbeck, Keely"]
```

```{python}
##need a function to take role skillz and people skillz and just return a single array

def return_consultant(full_name): 
   df = peopleSkillz[peopleSkillz['full_name'] == full_name].iloc[0,1:]
   return df

```

```{python}
return_consultant("Weisbeck, Keely")
return_consultant("Harmon, Paul")
```

```{python}

def return_role(role_name): 
   df = roleSkillz[roleSkillz['name'] == role_name].iloc[0,3:]
   return df

```

```{python}
return_role('Data Scientist')
```

```{python}

#skill: array of skills for a given person
#role: array of roles for a given person
def match_skillz(skill, role): 
  temp = skill >= role 
  pct = sum(temp)/temp.size
  return pct

```

```{python}
t1 = np.array([3,4,5])
t2 = np.array([4,2,2])
match_skillz(t1, t2)
```
```{python}
roleSkillz.iloc[0:3, 0:1]
```

```{python}

## Script
# pick a role
# pick a consultant

# run the comparison
personskill = return_consultant("Weisbeck, Keely") 
personskill
role = return_role('Data Scientist') 
role
temp = match_skillz(personskill, role)

print("Match Score is:\n", temp)

```