### 2021-09-21 Simulation with Single Profile.Rmd

---
title: "Simulation Study with Feature Contribution"
author: "Paul Harmon"
date: "9/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

source("OutlierCompNew.R")
library(ggplot2)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(janitor)
library(magrittr)
library(knitr)
library(Rtsne)
library(robustHD)
library(GGally)
library(stringr)
library(plotly)
library(R.utils)
library(vegan)
library(mnormt)

```

# Introduction

This doc looks at a profile where we have aberrant values only on a few of the features. Can we detect?

```{r, echo = FALSE, eval = TRUE}

n_sims = 100

#create a bunch of lists to store stuff later
plot_data_list <- list()
plot_tsne_results_list <- list()
plot_mds_results_list <- list()
plot_tsne_fstat_results_list <- list()
plot_mds_fstat_results_list <- list()
results_list <- list()
results_list_cmd <- list()
df3list = df2list = list()

```

## Profile

Let's look at a couple example profiles - here we might have a feature with only a few weird values.

```{r sims_v2, eval = TRUE, cache = TRUE, warning = FALSE, message = FALSE}

  #### Simulate the Data with outliers and a single group (simulates all from one group with mean 40)

  g1 <- rmnorm(n = 20, mean = rep(40, 10), diag(45, 10)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 20, mean = rep(40, 20), diag(45, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 20, mean = rep(40, 30), diag(45, 30)) %>%
    t() %>%
    data.frame()

  yes_structure <- rbind(g1, g2, g3) %>%
    data.frame() %>%
    mutate(Group = c(rep("1", 10), rep("2", 20), rep("3", 30)))

  yes_structure_scale <- lapply(yes_structure[,1:20], scale) %>% as_tibble()

  ### adds outlier - where it's roughly the same for all but 2 of the features
  add_outliers <- c(rnorm(18, 0, 1), c(-5,5))
  names(add_outliers) <- names(yes_structure_scale)

  yes_structure_final <- bind_rows(yes_structure_scale, add_outliers) 
  yes_structure_final$Group <- c(yes_structure$Group, rep("Outlier",1))

  ##### Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:20) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
  datplot

```

```{r}
perplexity = 10
  #### Run the Permanova mapping (for t-SNE)

  xTSNE <- MultiPermanova(yes_structure_final[,1:20], maptype = "tSNE")
  results_list[[j]] <- xTSNE

  df2 = tibble(Holdout = 1:length(xTSNE$modellist), PValues = sapply(xTSNE$modellist, pullPval), Fstat = sapply(xTSNE$modellist, pullFstat))
  df2$Colors <- ifelse(df2$PValues < 0.05, TRUE, FALSE)
  df2list[[j]] = df2

  #### TSNE Plots########################################################
  #Visualize the results and store in a list
  sigplot <- ggplot(df2, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: P-Values From Adonis")
  plot_tsne_results_list[[j]] <- sigplot

  #Visualize the f stats and store in a list
  sigplotf <- ggplot(df2, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: F Stats From Adonis")
  plot_tsne_fstat_results_list[[j]] <- sigplotf

   #### Run the Permanova mapping (for MDS)

   xCMD <- MultiPermanova(yes_structure_final[,1:10], maptype = "MDS")
  results_list_cmd[[j]] <- xCMD

  ### CMD Plots #########################################################

  df3 = tibble(Holdout = 1:length(xCMD$modellist), PValues = sapply(xCMD$modellist, pullPval), Fstat = sapply(xCMD$modellist, pullFstat))
    df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)
  df3list[[j]] = df3

  #Visualize the results and store in a list
  sigplot2 <- ggplot(df3, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: P-Values From Adonis")
  plot_mds_results_list[[j]] <- sigplot2

  #Visualize the f stats and store in a list
  sigplot2f <- ggplot(df3, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: F Stats From Adonis")
  plot_mds_fstat_results_list[[j]] <- sigplot2f

}